---
title: "R Basics"
author: "Ed Gonzalez"
date: "2023-01-03"
output:
  pdf_document: default
  html_document: default
---

```{r echo=FALSE}
install.packages("carData", repos = "http://cran.us.r-project.org")
library(carData)
```
```{r}
summary(TitanicSurvival)
```
# Using '?' to find information about a function and a quick method of creating a table with specific variables
```{r}
?table()
table(TitanicSurvival$passengerClass, TitanicSurvival$survived)
```

# Creating a histogram and boxplot using Base R functions
```{r}
hist(TitanicSurvival$age, 
     main = "Titanic Survival Histogram",
     xlab = "Age", 
     ylab = "Number of deaths")
```

* * *

```{r}
boxplot(age ~ survived, data = TitanicSurvival, 
        main = "Titanic Survival",
        xlab = "Age", 
        ylab = "Survived", horizontal = TRUE)
```

# Downloading a dataset and assigning variable names
```{r eval = TRUE}
abalone.url <- "https://archive.ics.uci.edu/ml/machine-learning-databases/abalone/abalone.data"
abalone.data <- read.csv(abalone.url, header=FALSE)
names(abalone.data) <- c("sex", "length", "diameter",
                         "height", "weight.whole", "weight.shucked",
                         "weight.viscera", "weight.shell", "rings")
```
# Functions that can identify the information contained within the data set, as well as creating variables
```{r}
class(abalone.data)
typeof("diameter")
summary(abalone.data$diameter[abalone.data$rings])
mean(abalone.data$diameter[abalone.data$sex == "F"])

x <- abalone.data$rings
x_t <- t(abalone.data$rings)
y <-abalone.data$diameter
```
# Doing some arithmetic with the variables we created
```{r}
solve(x_t%*%x)*(x_t%*%y)
```