---
title: "Joins"
author: "Ed Gonzalez"
output: pdf_document
---

```{r}
install.packages("tidyverse", repos = "http://cran.us.r-project.org")
library(tidyverse)

baseball01 <- read_csv("/Users/ed/Downloads/baseball01.csv")

baseball02 <- read_csv("/Users/ed/Downloads/baseball02.csv")
```

```{r}
ljbaseball <- left_join(baseball01, baseball02)
ljbaseball
```
With the left_join function, we get the above results. We now have a table with 
all of the Names from baseball01, but with the added variables in baseball02
with "NA" in cells that didn't contain any information.
```{r}
ijbaseball <- inner_join(baseball01, baseball02)
ijbaseball
```
Inner_join function grabs the Names that baseball01/02 have in common. In this
case, we have Tinker and Cobb as the only rows as they contain information for 
all of the variables in both data sets. 
```{r}
fjbaseball <- full_join(baseball01, baseball02)
fjbaseball
```
With full_join, it is fairly self-explanatory as it joins the data sets entirely
and inserts "NA" in any cells that doesn't contain data from either set.
