---
title: "Parallel Computing"
author: "Ed Gonzalez"
date: "2022-12-01"
output:
  pdf_document: default
  html_document: default
---
# I had to move onto the desktop version of R Studio for this as the cloud-based program did not allow me to use more than 0.5GB of RAM

```{r}
install.packages("readr", repos = "http://cran.us.r-project.org")
library(readr)

airlines09 <- read_csv("/Users/ed/Downloads/2009.csv")
```

# Installing some packages for tidying the data set and for parallel computing 
```{r}
install.packages("tidyverse", repos = "http://cran.us.r-project.org")
install.packages("sparklyr", repos = "http://cran.us.r-project.org")
install.packages("multidplyr", repos = "http://cran.us.r-project.org")

library(tidyverse)
library(sparklyr)
library(multidplyr)
library(dplyr)
```

# Timing some code using dplyr that finds the mean arrival delay of each carrier
```{r}
system.time(
  airline_delay <- airlines09 %>%
  group_by(OP_CARRIER) %>% 
  summarize(mean = mean(ARR_DELAY, na.rm = TRUE)) %>% 
    print())
```

# Same as before, but now utilizing the for %do% process and timing it
```{r}
library(foreach)

carriers<-split(airlines09,airlines09$OP_CARRIER)

system.time(
  {foreach(i=carriers,.combine=c) %do%
  mean(i$ARR_DELAY, na.rm = T) %>% 
  print()}
)
```

# Using the doParallel package where I essentially just change the %do% to %dopar% and assign some cores
```{r}
library(doParallel)
n.cores <- detectCores() - 1
registerDoParallel(n.cores)

system.time({
  foreach(i=carriers,.combine=c) %dopar%
  mean(i$ARR_DELAY, na.rm = T) %>% 
  print()}
)

```

# Same procee utilizng the multidplyr package
```{r}
library(multidplyr)
cl <- new_cluster(7)

system.time(
  {
  airline_delay <- airlines09 %>%
    group_by(OP_CARRIER) %>%
    partition(cl) %>% 
    summarize(mean = mean(ARR_DELAY, na.rm = TRUE)) %>% 
    collect() %>% 
    print()}
  )
```

# Creating a table with times for each computation
```{r}
tidyverse.time <- c(0.128, 0.045, 0.233)
foreach.do.time <- c(0.032, 0.021, 0.052)
foreach.dopar.time <- c(0.004, 0.025, 0.085)
multidplyr.time <- c(3.73, 1.834, 7.336)

rbind(tidyverse.time, foreach.do.time, foreach.dopar.time, multidplyr.time)
      

```

# Using dplyr to select for specific variables and performing a summary function to find the mean of arrival delay, counting unique instances, and finding a linear regression model between arrival delay and taxi in/out times
```{r}
system.time({
  taxi.times <- airlines09 %>% 
  select(ORIGIN, DEST, OP_CARRIER, ARR_DELAY, TAXI_IN, TAXI_OUT) %>% 
  drop_na() %>% 
  group_by(ORIGIN, DEST, OP_CARRIER) %>% 
  summarise(Count = n(),Mean = mean(ARR_DELAY, na.rm = T),
            fit = summary(lm(ARR_DELAY ~ TAXI_IN + TAXI_OUT))$r.squared)
  print(head(taxi.times))
})
```

# Doing the same as before but utilizing multidplyr functions to compare times
```{r}
system.time({
  taxi.times2 <- airlines09 %>% 
  select(ORIGIN, DEST, OP_CARRIER, ARR_DELAY, TAXI_IN, TAXI_OUT) %>% 
  drop_na() %>% 
  group_by(ORIGIN, DEST, OP_CARRIER) %>%
    partition(cl) %>% 
    summarise(Count = dplyr::n(),Mean = mean(ARR_DELAY, na.rm = T),
            fit = summary(lm(ARR_DELAY ~ TAXI_IN + TAXI_OUT))$r.squared)
  print(head(taxi.times2))
})
```

# Using ggplot2 package to create a scatter plot
```{r}
library(ggplot2)

scatter.p <- ggplot2::ggplot(taxi.times, aes(Count, fit))

scatter.p + geom_point() + theme_classic()
```
# Creating boxplots for each individual airline
```{r}
box.p <- ggplot2::ggplot(taxi.times, aes(OP_CARRIER, fit))

box.p+geom_boxplot(outlier.shape = NA ) + theme_classic()
```
