---
title: "Tidyverse and ggplot (cont.)"
author: "Ed Gonzalez"
output: pdf_document
---
```{r}
install.packages("tidyverse", repos = "http://cran.us.r-project.org")
install.packages("nycflights13", repos = "http://cran.us.r-project.org")

library(tidyverse)
library(lubridate)
library(nycflights13)

nyc <- nycflights13::flights
```

# A little messy, but here I explain what each part of the coding does using the dplyr functions and then creating a new variable within the code where I calculate the arrival 

```{r}
nycJan <- nyc %>% # this assigns the following code to a variable
  filter(month == 1) %>% # selecting only for the months that have "1" listed
  filter(dest == "ORD" | dest == "DSM" | dest == "STL" | dest =="MCI" | dest == "MDW") %>% # this allows us to only see data from the listed airports
  mutate(dep_time = dep_time/100) %>% #this transforms the times into a more recognizable format by adding a decimal
  unite("dep", c(year, month, day, dep_time), sep="/", remove = FALSE) %>% #unite is merging the variables listed and adding a slash for a separator 
  mutate(dep = ymd_hm(dep, tz = "America/New_York", quiet= TRUE)) %>% #this arranges the date and time in the designated format and also assigns a timezone
  filter(!is.na(dep))

nycJan
```


```{r}
nycJan <- nyc %>% 
  filter(month == 1) %>% 
  filter(dest == "ORD" | dest == "DSM" | dest == "STL" | dest == "MCI" | dest == "MDW") %>% 
  mutate(dep_time = dep_time/100) %>% 
  unite("dep", c(year, month, day, dep_time), sep ="/", remove = FALSE) %>% 
  mutate(dep = ymd_hm(dep, tz = "America/New_York", quiet = TRUE)) %>%
  filter(!is.na(dep)) %>% 
  mutate(arr_time = arr_time/100) %>% 
  unite("arr", c(year, month, day, arr_time), sep ="/", remove = FALSE) %>% 
  mutate(arr = ymd_hm(arr, tz = "America/New_York", quiet = TRUE)) %>%
  filter(!is.na(arr)) %>% 
  mutate(duration = dep - arr)
```
# This is an example of how I learned to use ggplot in a more concise way. Seeing a lot of personal growth 
```{r}
Jan_plot <- ggplot(nycJan, aes(air_time, duration))

Jan_plot + geom_point()
```
# A quick demo of the stringr package using everyone's favorite, Ames Housing data!
```{r}
install.packages("AmesHousing", repos = "http://cran.us.r-project.org")
library(AmesHousing)

Ames <- make_ordinal_ames()
AmesNeigh <-  fct_unique(Ames$Neighborhood)

str_count(AmesNeigh, "[B|b]")

str_count(AmesNeigh, "_")
```
# Demo of the purr package where I can create a boxplot with variables that have the word "Area" within it
```{r}
Ames %>% 
  select(ends_with("Area")) %>% 
  boxplot(outline = FALSE)
```