---
title: "Tidyverse & ggplot"
author: "Ed Gonzalez"
output: pdf_document
---

## I am using a dataset collected by a professor on my campus which recorded physical attributes about students

```{r, echo=TRUE}
install.packages("readr", repos = "http://cran.us.r-project.org")
install.packages("tidyverse", repos = "http://cran.us.r-project.org")

library(readr)
library(tidyverse)

KimData <- read_csv("Downloads/Clean-KimData.csv")
```

## Utilizing some dplyr functions here!

```{r}
# Demonstrating the | operator when filtering to display male or freshman students as well as the usage of %>%
FreshOrMales <-KimData %>% 
  filter(Semester < 2 | Gender == "M")

# Another demonstration of piping to create a tibble with selected variables 
KimDataPhysical <- KimData %>%
  select(Semester, Gender, `Shoe Size`, Height, Weight, Handed)
head(KimDataPhysical)

# Using mutate() to create new variables in the data set
KimDataPhysical <- KimDataPhysical %>% 
  mutate(Year = round(Semester/2))
head(KimDataPhysical)

KimDataPhysical <- KimDataPhysical %>% 
  mutate(BMI = 703 * (Weight/Height^2))
head(KimDataPhysical)

KimDataPhysical <- KimDataPhysical %>% 
  mutate(Obese = BMI >= 30)
head(KimDataPhysical)
```

## Using group_by() and summarize() functions

```{r}
KimDataPhysical %>% group_by(Year) %>%
  summarize( Year_BMI = mean(BMI, na.rm=TRUE))
```

## Finding the average shoe size, but then correcting for those (like me) that have a size 16 shoe

```{r}
shoe_size_Kim <- KimDataPhysical %>%
  summarize("Total number" = n(), "Average Shoe Size" = mean(`Shoe Size`, na.rm=TRUE))

shoe_size_Kim	
```

```{r}
regular_shoe_sizes <- KimDataPhysical %>% filter(`Shoe Size` < 16) %>% 
  summarize(count = n(), mean = mean(`Shoe Size`, na.rm = TRUE))

regular_shoe_sizes
```

# More dplyr! This time I'm poking around everyone's favorite data set, the Ames housing data set.

```{r}
install.packages("AmesHousing", repos = "http://cran.us.r-project.org")
install.packages("dplyr", repos = "http://cran.us.r-project.org")

library(AmesHousing)      
library(dplyr)

ames<-make_ames() 

Remodeled <- ames$Year_Built != ames$Year_Remod_Add

set.seed(248)
ames.500 <- sample_n(ames, 500)  
ames.500


ames.500$Remodeled <- ames.500$Year_Built != ames.500$Year_Remod_Add
ames.500$Remodeled

t.test(Sale_Price ~ Remodeled, data = ames.500)

plot(x = ames.500$Year_Built, y = ames.500$Year_Remod_Add)
```

## Running simulated data sets

```{r}
sample_1 <- rnorm(50,15,8)
sample_2 <- rnorm(50,17,8)

t.test(sample_1, sample_2)
```

## Increasing the sample size to see how it affects the p-value and our confidence in the results

```{r}
sample_1 <- rnorm(100,15,8)
sample_2 <- rnorm(100,17,8)

t.test(sample_1, sample_2)
```

```{r}
sample_1 <- rnorm(200,15,8)
sample_2 <- rnorm(200,17,8)

t.test(sample_1, sample_2)
```

## Finding the linear regression between the sale price and square footage and the year it was built.

```{r}
fit.original <- lm(Sale_Price ~ Year_Built + First_Flr_SF, data=ames)

fit.original

summary(fit.original)$adj.r.squared
```

```{r}
fit.updated <- lm(Sale_Price ~ Lot_Area + Overall_Qual, data=ames)

summary(fit.updated)$adj.r.squared

```

# How about some ggplot?

```{r}
ggplot(ames, aes(x = Sale_Price, y = Overall_Cond)) + geom_boxplot()
```

```{r}
ggplot(ames, aes(x = Overall_Cond, y = Sale_Price, color = Overall_Qual)) + geom_jitter( width = 0) + theme(axis.text.x = element_text(angle=30)) + labs(x = "Overall Condition of House", y = "Final Sale Price")
```

By adding a legend and color to the data points, we can see how the data is distributed in relation to its quality and sale price. This is where I first started to really learn ggplot. Later on I start using the piping method to better organize my code and improve replicability.
